\chapter{Current Copier Block (\tx{fanout} Block)}

The current copier block is an analog block available on the \hcdc
device~\cite{fan.h}. Figures~\ref{fan:values} and \ref{fan:types} presents a
complete summary of the digitally settable codes for the block. The current
copier accepts one analog input (\tx{in}) and produces three analog outputs (\tx{out0},
\tx{out1} and \tx{out2}), where the analog outputs are copies of the provided signal.

\noindent\textbf{Location}: Each slice of the \hcdc device contains two current
copier blocks. Given a slice at \hslice{chip}{tile}{slice}, the two current
copiers on the slice are written as \hindex{chip}{tile}{slice}{0}
and \hindex{chip}{tile}{slice}{1}.

\begin{marginfigure}
  \small
  \begin{tabular}{l|l}
    code &values\\
    \hline
    \tx{enable}& \tx{bool_t}\\ 
    \tx{third}& \tx{bool_t}\\ 
    \tx{range}& \tx{range_t}\\
    \tx{inv[out0Id]}& \tx{bool_t}\\
    \tx{inv[out1Id]}& \tx{bool_t}\\
    \tx{inv[out2Id]}& \tx{bool_t}\\
    \tx{nmos} & 8\\
    \tx{pmos}\caveat & 8\\
    \tx{port_cal[out0Id]}& 64\\
    \tx{port_cal[out1Id]}& 64\\
    \tx{port_cal[out2Id]}& 64\\
  \end{tabular}
  \caption{Fanout Values \cite{fu.h}}
  \label{fan:values}
\end{marginfigure}
  
\begin{marginfigure}
  \small
  \begin{tabular}{l|l}
    code & type \\
    \hline
    \tx{enable}& \static \\ 
    \tx{range}& \static \\
    \tx{inv[out0]}& \static \\
    \tx{inv[out1]}& \static \\
    \tx{inv[out2]}& \static \\
    \tx{nmos} & \hidden \\
    \tx{pmos}\caveat & \hidden \\
    \tx{port_cal[out0]}& \hidden \\
    \tx{port_cal[out1]}& \hidden\\
    \tx{port_cal[out2]}& \hidden \\
  \end{tabular}
  \caption{Fanout Code Types\cite{fu.h}}
  \label{fan:types}
\end{marginfigure}

  
\section{Block Function}\label{fanout:blockfun}
The behavior of output $i$ (\tx{out}$i$) the current copier is dictated by the
relation presented below. We write the analog input as $in$ in the presented
relation. The value returned by the function is the value of the current in $\mu
A$. Any behavior not covered by this algorithm is undefined.

\begin{algorithmic}
  \If{\tx{enable}}
  \State{sign(\tx{inv}[\tx{out}$i$]) \tx{in}} 
  \EndIf
\end{algorithmic}

The \tx{inv} code for the output $i$ determines the whether the copied signal
should be inverted or not. Note that all \static codes, with the exception of
the \tx{range} and \tx{third} codes, are used in the block function. The \tx{range} code is
used to configure the current limitations of the block, and the \tx{third} code
determines if the third output (\tx{out2}) of the current copier is in use.

\subsection{Operating Ranges}

The magnitude of the analog input $in$ must fall within the current limits of
the current copier. These limits are determined by the \tx{range} code:

\begin{algorithmic}
  \State{$\tx{in} \in scale(\tx{range}) \cdot [-2 \mu A, 2 \mu A] $}
  \State{$\tx{out0} \in scale(\tx{range}) \cdot [-2 \mu A, 2 \mu A] $}
  \State{$\tx{out1} \in scale(\tx{range}) \cdot [-2 \mu A, 2 \mu A] $}
  \State{$\tx{out2} \in scale(\tx{range}) \cdot [-2 \mu A, 2 \mu A] $}
\end{algorithmic}
\section{Calibration}\label{fanout:calib}

The fanout block has five hidden codes:

\begin{itemize}
\item\tx{port_cal[out0]}, \tx{port_cal[out1]}, \tx{port_cal[out2]}: These bias
  correction codes control the currents injected into
  the \tx{out0}, \tx{out1} and \tx{out2} outputs of the current copier. These
  injected currents are used to correct for any unwanted biases in the block. A
  \tx{port_cal} value of 32 approximately corresponds to an injected current of zero.
\item\tx{pmos} and \tx{nmos}: These current reference codes affect the magnitude of the
  bias correction codes. They correspond to iref currents in the schematic of
  the block. 
\end{itemize}

The \tx{pmos} code is always set to $3$ and the \tx{nmos} code is always set to
$0$. The remaining codes are set by the block's calibration routine. The calibration routine implements the following algorithm:

\begin{algorithmic}
  \State {ctbl = \tx{make_table()}}
  \State {tbl0 = \tx{make_table()}}
  \For{cal0 in 0...64 with stride=4}
  \For{cal1 in 0...64 with stride=4}
  \For{cal2 in 0...64 with stride=4}
  \State {loss = obj(cal0,cal1,cal2)}
  \State {ctbl $\leftarrow$ loss,(cal0,cal1,cal2)}
  \EndFor
  \EndFor
  \EndFor
  \State {tbl = \tx{make_table()}}
  \For{i in -3..3}
  \State{cal0 = ctbl.cal0+i}
  \For{j in -3..3}
  \State{cal1 = ctbl.cal1+j}
  \For{k in -3..3}
  \State{cal2 = ctbl.cal2+k}
  \State {loss = obj(cal0,cal1,cal2)}
  \State{tbl $\leftarrow$ loss,(cal0,cal1,cal2)}
  \EndFor
  \EndFor
  \EndFor
  \State{return tbl.cal0,tbl.cal1,tbl.cal2}
\end{algorithmic}

At a high level, the calibration algorithm independently finds the best
coarse-grained assignments for the \tx{port_cal[out0Id]}, \tx{port_cal[out1Id]}
and \tx{port_cal[out2Id]} bias correction codes. It then fine tunes the
coarse-grained set of assignments by doing a local search in the vicinity of values
around the coarse-grained set of assignments. 

The objective function used by the fanout calibration routine  XXX


\section{Profiling}\label{fanout:calib}


\section{Grendel API Hook}

\grendel supports configuring fanout blocks using the \tx{use_fanout} command.
We present the general formulation of the \tx{use_fanout} command below:

\lstset{ 
  morekeywords={use_fanout,src,sgn,val,rng,three,two},
  basicstyle=\small
}
\begin{lstlisting}
use_fanout chip tile slice index sgn sign0 sign1 sign2 rng range three|two
\end{lstlisting}

The \tx{use_fanout} command accepts a fanout location in the form of a chip,
tile, slice and index and several additional arguments
(\tx{sign0}, \tx{sign1}, \tx{sign2}, \tx{range}, \tx{three}, and \tx{two}) that set
the \static and \dynamic codes of the block:
\begin{itemize}
\item \tx{sign0},\tx{sign1} and \tx{sign2} arguments: The sign arguments set the
  \tx{inv} static codes for the first, second and third outputs respectively.
  These arguments accept \tx{+} and \tx{-} as values. If the sign argument is
  \tx{-}, the corresponding \tx{inv} code is set to true.
  \item\tx{range} argument: The range argument sets the \tx{range} static code
    of the block. This argument accepts \tx{m} and \tx{h} as values, where
    \tx{m} corresponds to \tx{RANGE_MED} and \tx{h} corresponds to
    \tx{RANGE_HIGH}.
    \item\tx{three} or \tx{two} argument: This argument determines if the third
      output (\tx{out2}) is enabled. If this argument is set to \tx{three}, the
      third output is enabled.
  \end{itemize}

\subsection{Example Usage}

The following invocation configures the fanout on chip 0, tile 0, slice 0, index
0 to copy the input signal, where the input signal is within $[-2 \mu A, 2 \mu
A]$. The first and second outputs are negated, and the third output is enabled.

\begin{lstlisting}
use_fanout 0 0 0 0 sgn - - + rng m three
\end{lstlisting}

The following invocation configures the fanout on chip 1, tile 0, slice 2, index
1 to copy the input signal, where the input signal is within $[-20 \mu A, 20 \mu
A]$. The second output is negated, and the third output is disabled.

\begin{lstlisting}
use_fanout 1 0 2 1 sgn + - + rng h two 
\end{lstlisting}